%%% Please fill in basic information on your thesis, which will be automatically
%%% inserted at the right places. You need to replace \xxx{...} by real data.

% Type of your thesis:
%	"bc" for Bachelor's
%	"mgr" for Master's
%	"phd" for PhD
%	"rig" for rigorosum
\def\ThesisType{bc}

% Language of your study programme:
%	"cs" for Czech
%	"en" for English
\def\StudyLanguage{en}

% Thesis title in English (exactly as in the official assignment)
% (Note: \xxx is a "ToDo label" which makes the unfilled visible. Remove it.)
\def\ThesisTitle{Kempe chains and rooted minors}

% Author of the thesis (you)
\def\ThesisAuthor{Tigran Arsenyan}

% Year when the thesis is submitted
\def\YearSubmitted{2025}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% see https://www.mff.cuni.cz/en/faculty/organizational-structure,
% or a full name of a department outside MFF)
\def\Department{Computer Science Institute of Charles University}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{institute}

% Thesis supervisor: name, surname and titles
\def\Supervisor{prof. Mgr. Zdeněk Dvořák, Ph.D.}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{Computer Science Institute of Charles University}

% Study programme (does not apply to rigorosum theses)
\def\StudyProgramme{Computer Science}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, who provided you with tea and pizza, etc.)
\def\Dedication{%
First, I would like to thank my supervisor, Prof. Zdeněk Dvořák, for his endless patience and invaluable guidance, as well as for pointing out the mistakes I made along the way.
I am also deeply grateful to my family for their support throughout my studies, and to my friends and loved ones for always being there for me.
Finally, I would like to acknowledge the contributions of the authors whose works I have cited in this thesis.
}


% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
Kempe chains are one of the basic tools used to prove variations of Hadwiger's conjecture. 
Thus, it is important to understand what the existence of particular Kempe chains guarantees in terms of (rooted) minors. 
In this thesis, we survey the known results on this topic and focus on the results of Kriesell and Mohr, who studied the following question: 
For any transversal $T$ of a coloring $\mathfrak{C}$ of order $k$ of a graph $G$, such that any pair of color classes induces a connected subgraph, 
does there exist a partition $H$ of a subset of $V(G)$ into connected sets such that $T$ is a transversal of $H$ and the sets of $H$ are 
pairwise adjacent whenever the corresponding vertices in $T$ are connected by a 2-colored path? This is open for $k \geq 5$; 
they proved the case $k = 5$ when the subgraph induced by $T$ is connected. They also showed that for $k \geq 7$, it is not sufficient to only 
guarantee 2-colored paths. We contribute by using computer-assisted enumeration to search for counterexamples for $k=6$, and show that minimal counterexamples $G$
for $k=5$ must be 2-connected.
}


% 3 to 5 keywords (recommended) separated by \sep
% Keywords are useful for indexing and searching for the theses by topic.
\def\ThesisKeywords{%
Kempe Chain\sep Rooted Minor \sep Hadwiger's Conjecture
}

% If any of your metadata strings contains TeX macros, you need to provide
% a plain-text version for use in XMP metadata embedded in the output PDF file.
% If you are not sure, check the generated thesis.xmpdata file.
\def\ThesisAuthorXMP{\ThesisAuthor}
\def\ThesisTitleXMP{\ThesisTitle}
\def\ThesisKeywordsXMP{\ThesisKeywords}
\def\AbstractXMP{%
Kempe chains are important tools in proofs related to Hadwiger's conjecture. This thesis surveys known results, focusing on a question by Kriesell and Mohr regarding colorings with connected color classes and the existence of routing graphs. For k = 6, no counterexamples were found in graphs up to 13 vertices. We also show that any minimal counterexample for k = 5 must be 2-connected.
}

% If your abstracts are long and do not fit in the infopage, you can make the
% fonts a bit smaller by this setting. (Also, you should try to compress your abstract more.)
\def\InfoPageFont{}
%\def\InfoPageFont{\small}  % uncomment to decrease font size

% If you are studing in a Czech programme, you also need to provide metadata in Czech:
% (in English programmes, this is not used anywhere)

\def\ThesisTitleCS{\xxx{Název práce česky}}
\def\DepartmentCS{\xxx{Název katedry česky}}
\def\DeptTypeCS{\xxx{Katedra}}
\def\SupervisorsDepartmentCS{\xxx{katedra vedoucího}}
\def\StudyProgrammeCS{\xxx{studijní program}}

\def\ThesisKeywordsCS{%
\xxx{klíčová slova\sep klíčové fráze}
}

\def\AbstractCS{%
\xxx{Abstrakt práce přeložte také do češtiny.}
}
